---
title: "Penguins size comparisons"
author: "Will Geiken"
date: "11/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(car)
library(ggbeeswarm)
library(effsize)
library(broom)
library(tidyverse)
```

# Part 1. T-tests

We use t-tests to compare means between 2 dif groups.

Mass comparison: male & female chinstrap penguins

```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap")

# Histograms
ggplot(data = chinstraps, aes(x = body_mass_g)) +
  geom_histogram(bins = 10) +
  facet_wrap(~sex)

#QQ plot (compares our sample to data if it were drawn from a perfectly normally distributed data set)

ggplot(data = chinstraps, aes(sample = body_mass_g))+
  geom_qq() +
  facet_wrap(~sex)

# Get descriptive statistics

chinstraps %>% 
  group_by(sex) %>% 
  summarize(
    mean_mass = mean(body_mass_g),
    median_mass = median(body_mass_g),
    sd_mass = sd(body_mass_g),
    sample_size = n())
```

Get vectors of body_mass_g for female and male chinstrap penguins separately

```{r}
chinstraps_f <- chinstraps %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g)

chinstraps_m <- chinstraps %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)
```

### T-tests, a couple different ways

1. Is there a significant difference in body mass between male and female chinstrap penguins?

```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)

chinstrap_ttest

cohen.d(chinstraps_f, chinstraps_m)
#gives a magnitude of 1.26 which is considered large, meaning it's a clear difference between female and male mass
```

Are male chinstrap penguins significantly larger (body mass) than female chinstrap penguins?

This means we're concerned with only one tail of the distribution

```{r}
t.test(chinstraps_m, chinstraps_f, alternative = "greater")

#can add "greater" or "less" depending on which vector is listed first

#called a one-sided t-test
```

