---
title: "Penguins size comparisons"
author: "Will Geiken"
date: "11/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(car)
library(ggbeeswarm)
library(effsize)
library(broom)
library(tidyverse)
```

# Part 1. T-tests

We use t-tests to compare means between 2 dif groups.

Mass comparison: male & female chinstrap penguins

```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap")

# Histograms
ggplot(data = chinstraps, aes(x = body_mass_g)) +
  geom_histogram(bins = 10) +
  facet_wrap(~sex)

#QQ plot (compares our sample to data if it were drawn from a perfectly normally distributed data set)

ggplot(data = chinstraps, aes(sample = body_mass_g))+
  geom_qq() +
  facet_wrap(~sex)

# Get descriptive statistics

chinstraps %>% 
  group_by(sex) %>% 
  summarize(
    mean_mass = mean(body_mass_g),
    median_mass = median(body_mass_g),
    sd_mass = sd(body_mass_g),
    sample_size = n())
```

Get vectors of body_mass_g for female and male chinstrap penguins separately

```{r}
chinstraps_f <- chinstraps %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g)

chinstraps_m <- chinstraps %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)
```

### T-tests, a couple different ways

1. Is there a significant difference in body mass between male and female chinstrap penguins?

```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)

chinstrap_ttest

cohen.d(chinstraps_f, chinstraps_m)
#gives a magnitude of 1.26 which is considered large, meaning it's a clear difference between female and male mass
```

Are male chinstrap penguins significantly larger (body mass) than female chinstrap penguins?

This means we're concerned with only one tail of the distribution

```{r}
t.test(chinstraps_m, chinstraps_f, alternative = "greater")

#can add "greater" or "less" depending on which vector is listed first

#called a one-sided t-test
```

## One-way ANOVA

Compares means across more than 2 groups.

## 1. Table of counts by species and sex 

```{r}
penguin_counts <- penguins %>% 
  count(species, sex)

penguin_counts
```

Create a subset of only female penguins

```{r}
penguin_f <- penguins %>% 
  filter(sex == "female")

ggplot(data = penguin_f, aes(x = flipper_length_mm)) +
  geom_histogram(aes(fill = species)) +
  facet_wrap(~species, scales = "free")

#scales = free allows for individualized x and y axes

ggplot(data = penguin_f, aes(sample = flipper_length_mm)) +
  geom_qq(aes(color = species)) +
  facet_wrap(~species, scales = "free")
```

### Find summary statistics and check for equal variances

Starting from penguins_f, create a summary table that groups by species, and returns the following:

-mean of flipper length
-sd "
-sample size
-variance 'var'
-standard error = sd/sqrt(n)

```{r}
penguins_summary <- penguin_f %>% 
  group_by(species) %>% 
  summarize(
    mean_flipper_length = mean(flipper_length_mm),
    sd_flipper_length = sd(flipper_length_mm),
    sample_size = n(),
    se_flipper_length = sd(flipper_length_mm)/sqrt(n()),
    var_flipper_length = var(flipper_length_mm))

#if largest variance is <4x your smallest variance, you're chilling and can run an ANOVA

#Levene's test is cautionary because the larger your sample, the more likely your result is significant

leveneTest(flipper_length_mm ~ species, data = penguin_f)
#looks at variance in flipper length as a function of species

#p > 0.05, so we retain our null hypothesis that the variances are close enough
```

### Beeswarm plot with mean +/- 1 sd

```{r}
ggplot() +
  geom_beeswarm(data = penguin_f, 
                aes(x = species, y = flipper_length_mm),
                color = "gray70",
                size = 3,
                alpha = 0.5) +
  geom_point(data = penguins_summary, aes(x = species, y = mean_flipper_length),
             size = 4) +
  geom_errorbar(data = penguins_summary,
                aes(x = species, 
                    ymax = mean_flipper_length + sd_flipper_length,
                    ymin = mean_flipper_length - sd_flipper_length),
                width = 0.1) +
  theme_light()
                
#IF it's refering to a variable, it goes inside aes, if not, it goes outside.
```

### One-way ANOVA

aov() by default runs a balanced ANOVA, usually if you have unequal sample sizes, you want to do a type 3 anova. We're doing both.

```{r}
penguin_aov <-
  aov(flipper_length_mm ~ species, data = penguin_f)

summary(penguin_aov)
#F value is huge, so the variance between groups is very large, so P is small, so we can reject null that the means for all of these groups for the three penguin types are the same.

#But we don't know what's different, so we need a post-hoc test

TukeyHSD(penguin_aov)
  #shows pairwise adjusted p-values. All are below 0.05, so it's likely that they are all from different populations

#If we have very unbalanced design, you probably want to do an orthoganol/type 3 anova
```

Type III Anova with the car package:
```{r}
penguin_lm <- lm(flipper_length_mm ~ species, data = penguin_f)
penguin_aov_t3 <- Anova(penguin_lm, type = "III")
penguin_aov_t3
#in this case, we had balanced design/similar sample sizes, so it gave us a similar result to the anova above
```

